A missing guardrail: The failure of protest risk to deter election manipulation in non-democracies

### 2023-06-09 Output/Submission files/manuscript_anon_v2.Rmd
Beginning cleaning up of appendices

There are thought to be two mechanisms by which election manipulation leads to protest, and thereby might restrain manipulation. The first is through revealing information about the strength of the government or ruling party. This information, in turn, feeds into strategic decision-making by opposition groups.  This logic is often seen in formal models [@Egorov2021;@chernykh2015third]. Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @fearon_self-enforcing_2011;@Luo2020;@Rozenas2016b]. The risk of mass protest in these models is assumed to constitute a significant guardrail for election integrity in democracies, and a significant challenge for ruling parties in non-democracies. Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014; @daxecker2012], civil society groups [@birch2017getting], and courts [@Chernykh2015]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance. Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons and across time\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.

@@

The literature on protest as a deterrent to manipulation also often overlooks the functions that elections, even highly controlled elections, may serve in non-democratic settings. For instance, Simpser [-@Simpser2011] shows that high levels of manipulation can actually induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength. Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with costlier forms of manipulation associated with less protest [@Harvey2020]. In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006]. Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation. In sum, election manipulation can be performed by strong regimes for reasons that go beyond 'mere winning.'

@@

The strategic and social-psychological explanations are, of course, interconnected. As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something. Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304]. Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014a;@Harvey2020;@Rod2019].

@@

@@

These hypotheses are tested using data from three datasets: NELDA [@hyde2012], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA. This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.  I limit the sample by excluding closed autocracies (where no national elections are held). Since the theory of popular backlash to manipulation has largely been articulated in an electoral-democratic context, I exclude liberal democracies from the main analysis.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 or 2.] As a robustness check for this sampling decision, I also run the main models with data including liberal democracies in the appendix, with no substantial changes in the outcomes. The first set of models, focusing on Hypothesis 1, include all the remaining cases in the dataset. For the subsequent models, focusing on post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud. Finally, the beginning of the date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012. The resulting dataset includes 1110 total election-period observations across 118 countries.

@@

```{r, include=F}
incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
                                            'Legislative constraints' = legcon.1lag, 
                                            'Alternative information' = altinf.1lag,
                                          'Liberal democracy index' = libdem.1lag,
                                            'GDP growth rate' = gdpgro.1lag, 
                                            'Civil society openness' = osorg.1lag, 
                                            'Public sector corruption' = pubseccorrup.lag,
                                            'National party organization' = natlpartyorg.lag, 
                                             'Election fraud' = v2elirreg.inv, 
                                            'Presidential election' = presidential,
                                            'Incumbent term-limited' = inc.termlimit.num,
                                            'Physical integrity index' = physinteg.1lag,
                                            'Urbanization' = urban.pct, 
                                            'Rentier state' = rentierstate,
                                            'Tax state' =  taxstate,
                                            'Winning vote-share' = winner.share.upd, 
                                            'Number of pre-election protests' = n.events.pre,
                                            'Off-schedule election' = unscheduled_election,
                                          'Regional protest diffusion' = regional_success_lag)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))

@@

```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Summary statistics"}
library(kableExtra)
datasummary_skim(incl.vars, output = "kableExtra")  %>%
    kable_styling(font_size = 9, full_width = FALSE)
```
\doublespace

@@

Finally, several binary control variables are included. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@Simpser2011] and a greater risk of protest. Next, a variable from NELDA indicates if the chief executive is *term limited* during the election at hand. This variable is included since incumbents running up against constitutional term limits can pose major challenges for the unity of the regime, possibly leading to fractured manipulation efforts and post-election protest [@hale_patronal_2014]. Finally, I include dummy variables that provide a categorical evaluation of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. States that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018]. Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.   Figure \@ref(fig:summary-table) shows summary statistics for all of these variables. To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.

@@

modelsummary(c(m1, m1b, m1c, m1a, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Random-intercept multilevel model of election-day fraud', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election")) %>% kable_styling(latex_options = "HOLD_position")

@@

```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.novoteshare.hurd.mm)

model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.base.hurd.mm)

model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag + pubseccorrup.lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.n.base2.hurd.mm)

model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.griev.hurd.mm)

model.n.griev2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.griev2.hurd.mm)

model.n.griev3.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag  ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.griev3.hurd.mm)

model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.repr.hurd)
```

```{r models-hurdle-size, warning = FALSE, message = FALSE, include = F}
model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.novoteshare)

model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.base)

model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.base2)

model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.griev)

@@

```{r zero-model-table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

@@

t1a <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 3: Zero-inflated models of protest initiation")

@@

t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "libdem.1lag", "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")

theme_compact(t1b, header_rows = TRUE, header_cols = FALSE)

models_t2 <- list()
models_t2[['Model 6']] <- model.n.base.hurd.mm
models_t2[['Model 7']] <- model.n.base2.hurd.mm
models_t2[['Model 8']] <- model.n.griev.hurd.mm
models_t2[['Model 9']] <- model.size.base
models_t2[['Model 10']] <- model.size.base2
models_t2[['Model 11']] <- model.size.griev

t2a <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("rentierstate1", "taxstate1", "n.events.pre", "unscheduled_election1", "inc.lose1", "v2elirreg.inv:winner.share.upd", "gdpgro.1lag:v2elirreg.inv"), statistics = character(0)) %>% set_caption("Table 4: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)

@@

t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), omit_coefs = c("(Intercept)", "jucon.1lag" , "legcon.1lag" , "gdpgro.1lag" , "altinf.1lag" , "libdem.1lag",  "pubseccorrup.lag" , "osorg.1lag" , "econ.crisis.nelda.1lag1" , "natlpartyorg.lag" , "v2elirreg.inv" , "pubseccorrup.lag"   , "presidential1" , "inc.termlimit.num1"  ,  "physinteg.1lag" , "physintlag.sq" , "urban.pct")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)

>>>>

There are thought to be two mechanisms by which election manipulation leads to protest, and thereby might restrain manipulation. The first is through revealing information about the strength of the government or ruling party. This information, in turn, feeds into strategic decision-making by opposition groups.  This logic is often seen in formal models [@Egorov2021;@chernykh2015third]. Several prominent models of electoral manipulation argue that the risk of mass post-election protest helps restrain incumbents who might otherwise prefer larger levels of manipulation [@magaloni2010; @Little2015; @Little2012; @birch2011; @fearon_self-enforcing_2011;@Luo2020;@Rozenas2016b]. The risk of mass protest in these models is assumed to constitute a significant guardrail for election integrity in democracies, and a significant challenge for ruling parties in non-democracies. Building on this logic, researchers have proposed that third-party institutions---like election monitors [@Little2012;@Hyde2014;@daxecker2012], civil society groups [@birch2017getting], and courts [@chernykh2015third]---can reduce manipulation and increase protest risk by revealing information about electoral malfeasance. Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons and across time\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.

@@

The literature on protest as a deterrent to manipulation also often overlooks the functions that elections, even highly controlled elections, may serve in non-democratic settings. For instance, Simpser [-@simpser2013] shows that high levels of manipulation can actually induce cooperation from opposition groups, elites, and other actors by revealing information about regime strength. Similarly, the type of manipulation employed can convey information about the ruling party's capacity to mobilize, with costlier forms of manipulation associated with less protest [@Harvey2020]. In this way, large-scale manipulation can contribute to a ruling party's ability to produce a large, stabilizing majority vote-share [@magaloni2006]. Similarly, Luo and Rozenas [-@Luo2020] posit that ex post fraud is more likely to result in protest than ex ante manipulation. In sum, election manipulation can be performed by strong regimes for reasons that go beyond 'mere winning.'

@@

The strategic and social-psychological explanations are, of course, interconnected. As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something. Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304]. Several empirical studies have, accordingly, found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019].

@@

## Set max_size zero for elections with no protest

nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

@@

These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA. This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections.  I limit the sample by excluding closed autocracies (where no national elections are held). Since the theory of popular backlash to manipulation has largely been articulated in an electoral-democratic context, I exclude liberal democracies from the main analysis.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 or 2.] As a robustness check for this sampling decision, I also run the main models with data including liberal democracies in the appendix, with no substantial changes in the outcomes. The first set of models, focusing on Hypothesis 1, include all the remaining cases in the dataset. For the subsequent models, focusing on post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud. Finally, the beginning of the date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012. The resulting dataset includes 1110 total election-period observations across 118 countries.

@@

```{r, include=F}
incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
                                            'Legislative constraints' = legcon.1lag, 
                                            'Alternative information' = altinf.1lag,
                                          'Liberal democracy index' = libdem.1lag,
                                            'GDP growth rate' = gdpgro.1lag, 
                                            'Civil society openness' = osorg.1lag, 
                                            #'Public sector corruption' = pubseccorrup.lag,
                                            'National party organization' = natlpartyorg.lag, 
                                             'Election fraud' = v2elirreg.inv, 
                                            'Presidential election' = presidential,
                                            'Incumbent term-limited' = inc.termlimit.num,
                                            'Physical integrity index' = physinteg.1lag,
                                            'Urbanization' = urban.pct, 
                                            'Rentier state' = rentierstate,
                                            'Tax state' =  taxstate,
                                            'Winning vote-share' = winner.share.upd, 
                                            'Number of pre-election protests' = n.events.pre,
                                            'Off-schedule election' = unscheduled_election,
                                          'Regional protest diffusion' = regional_success_lag)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))

@@

```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)
datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics")  %>%
    kable_styling(font_size = 9, full_width = FALSE)
```
\doublespace

@@

Finally, several binary control variables are included. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest. Next, a variable from NELDA indicates if the chief executive is *term limited* during the election at hand. This variable is included since incumbents running up against constitutional term limits can pose major challenges for the unity of the regime, possibly leading to fractured manipulation efforts and post-election protest [@hale_patronal_2014]. Finally, I include dummy variables that provide a categorical evaluation of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. States that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018]. Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.   Table \@ref(tab:summary-table) shows summary statistics for all of these variables. To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.

@@

modelsummary(c(m1, m1b, m1c, m1a, m1d), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion"
                 )) %>% kable_styling(latex_options = "HOLD_position")

@@

```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
model.n.novoteshare.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.novoteshare.hurd.mm)

model.n.base.hurd.mm <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.base.hurd.mm)

model.n.base2.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.n.base2.hurd.mm)

model.n.griev.hurd.mm <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  v2elirreg.inv*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.griev.hurd.mm)

model.n.repr.hurd <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +   + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv  +  physinteg.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.repr.hurd)
```

```{r models-hurdle-size, warning = FALSE, message = FALSE, include = F}
model.size.novoteshare <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.novoteshare)

model.size.base <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   v2elirreg.inv + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.base)

model.size.base2 <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2elirreg.inv +  +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.base2)

model.size.griev <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  v2elirreg.inv*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.size.griev)

@@

```{r initiation-table-setup, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

@@

t1a <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Intercept" = "(Intercept)", "Judicial ind. (lag)" = "jucon.1lag", "Liberal democracy index" = "libdem.1lag",
            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info. (lag)" = "altinf.1lag",                             "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag",                "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Presidential" = "presidential1", "Incumbent term-limited" = "inc.termlimit.num",
            "Physical integrity (lag)" = "physinteg.1lag", "Phys. int. (lag) squared" = "physintlag.sq",
            "Urban" = "urban.pct", "Rentier state" = "rentierstate1", "Tax state" = "taxstate1"),
              statistics = character(0))  %>% set_caption("Table 3: Zero-inflated models of protest initiation")

@@

t1b <- huxreg(models_t1, tidy_args = c("component" = "zi"), borders = 0, number_format = 2, coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 3 cont.: Zero-inflated models of protest initiation")

theme_compact(t1b, header_rows = TRUE, header_cols = FALSE)
```

```{r conditional-table, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
models_t2 <- list()
models_t2[['Model 6']] <- model.n.base.hurd.mm
models_t2[['Model 7']] <- model.n.base2.hurd.mm
models_t2[['Model 8']] <- model.n.griev.hurd.mm
models_t2[['Model 9']] <- model.size.base
models_t2[['Model 10']] <- model.size.base2
models_t2[['Model 11']] <- model.size.griev

t2a <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), coefs = c("Intercept" = "(Intercept)", "Judicial ind. (lag)" = "jucon.1lag",
            "Legislative constraints (lag)" = "legcon.1lag", "Alternative info. (lag)" = "altinf.1lag",                             "Liberal democracy index (lag)" = "libdem.1lag", "GDP growth rate (lag)" = "gdpgro.1lag",                               "Civil society openness (lag)" = "osorg.1lag", "National party organization (lag)" = "natlpartyorg.lag",                "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Presidential" = "presidential1", "Incumbent term-limited" = "inc.termlimit.num",
            "Physical integrity (lag)" = "physinteg.1lag", "Phys. int. (lag) squared" = "physintlag.sq",
            "Urban" = "urban.pct", "Rentier state" = "rentierstate1", "Tax state" = "taxstate1"), statistics = character(0)) %>% set_caption("Table 4: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)

@@

t2b <- huxreg(models_t2, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, stars = c(`***` = 0.001, `**` = 0.01, `*` = 0.05, `+` = .1), coefs = c("Number of pre-election protests" = "n.events.pre", "Off-schedule election" = "unscheduled_election1", "Regional protest diffusion" = "regional_success_lag", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd", "Election fraud : GDP growth (lag)" = "gdpgro.1lag:v2elirreg.inv")) %>% set_caption("Table 4 cont.: Negative binomial models of number and duration of protests") %>% insert_row("", "", "Number", "", "", "Size", "", after = 0)
